---
title: "Background"
author: "Mitchell Stares"
weight: 36
date: "2018-11-11"
output: blogdown::html_page
editor_options:
  chunk_output_type: console
---
```{r knitr options, message=FALSE, warning=FALSE, include=FALSE}
library(knitr)
knitr::opts_chunk$set(warning = FALSE, message = FALSE)

```
```{r message=FALSE, warning=FALSE, include=FALSE}
library(tidyverse)
iris.box <- ggplot(iris, aes(x=Species,y=Sepal.Length)) +
  geom_boxplot()
iris.hist <- ggplot(iris, aes(x=Sepal.Length)) +
  geom_histogram()
iris.scatter <- ggplot(iris, aes(x=Sepal.Length, y=Petal.Length, colour=Species)) +
  geom_point()
```

The plot and legend background colours can be changed using the following:  

Theme argument                                       | Description  
-----------------------------------------------------|---------------------------------------------  
panel.background = element_rect(insert changes here) |This changes the background of the main plot itself. We need element_rect() as it is a rectangle geometric object.   
legend.background = element_rect(insert changes here) |This will change the main area of the legend.  
legend.key = element_rect(insert changes here) |This will change the small boxes that each of the factors levels are identified with.  

For all arguments, you can replace the element_rect(), element_line() etc. with element_blank() to remove it.  

Within each of the element_rect() we can change various things. The most common ones are:

Element argument | Description  
-----------------|--------------------------
fill = "colour" | This will change the overall colour of the object.  
colour = "colour" | This will change the outline of the rectangle.  
size = number | This will change the size/thickness of font and lines.  

Each of the "colour" arguments can be a specified a number of ways. The most common way is using one of the MANY predefined colours within R. A quick run down of these can be found [here](http://www.stat.columbia.edu/~tzheng/files/Rcolor.pdf). For any of these, just put the name as it is spelt in that guide in quotations.  
```{r}
iris.scatter + theme(panel.background = element_rect(fill="lavender", colour="red"), legend.background = element_rect(fill="lavender", colour="yellow", size=1), legend.key = element_rect(fill = "gray50", colour = "green", size = 0.5)) 
```
